---
title: "exercise 2"
author: "ajay parihar"
date: "April 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Exercise 2

The data set Houses shows the yearly number of family houses sold in Belgium
(Houses) and the total price of transactions (Houseprices) from 1973 up to 2017.
First, create the average house price by dividing the total price of transactions by
the yearly number of family houses sold. This new variable will be analysed.
Split the data in a training set up to 2010 and a test set from 2011 up to 2017.
Use the training set for estimation of the methods/models, and use the test set for
assessing the forecast accuracy.
In each step of the exercise, discuss your results and explain your choices. Use
additional tables and graphs wherever they clarify your answer.

```{r }


library(fpp2)
library(readxl)
library(portes)
#install.packages("lmtest")
library(lmtest)

setwd("C:/Users/aparihar/Downloads")

# read the data
houses <- read_excel("DataSets.xlsx", sheet = "Houses")
houses$averageprice<-houses$Houseprices/houses$Houses
str(houses)
###create time series
rsvhouse <- ts(houses[,4], frequency = 1, start = 1973)
rsv1house <- window(rsvhouse, end=2009)
rsv2house <- window(rsvhouse, start=2010)
h=length(rsv2house)
h
```
##exercise2.1
Explore the data using relevant graphs.

here we plot the average price graph for the time period, we don't see any seasonality in the time series

```{r }
# Plot the data
plot(rsvhouse)
lines(rsv1house, col="red")
lines(rsv2house, col="blue")


```


##Exercise2.2
 Create forecasts using the most appropriate naive method. Check the residual
diagnostics and the forecast accuracy.
We run the naive method, a plot is created. Looking at the residuls we don't see any white noise

```{r }
##here we run naive method 
f2 <- naive(rsv1house, h=h)
accuracy(f2,rsv2house)

##lets plot the grpah
plot(rsvhouse,main="avg house price", ylab="",xlab="Day")
lines(f2$mean,col=2)
legend("bottomleft",lty=1,col=2,legend="Naive Method")
#now we check residuals
checkresiduals(f2)




```

## Exercise 2.3
Q)Generate forecasts using the relevant exponential smoothing methods. Check
the residual diagnostics and the forecast accuracy
A) h2 damped is the best 

```{r }
h1 <- holt(rsv1house, h=100)

h2 <- holt(rsv1house, h=100, damped=TRUE)

h3 <- holt(rsv1house, h=100, exponential=TRUE, damped=TRUE)

plot(h1, type="l", ylab="Average House Prices",

xlab="Year", fcol="white", shadecols="white")
lines(fitted(h1), col=2)

lines(fitted(h2), col=3)

lines(fitted(h3), col=4)



lines(h1$mean, col=2, type="l")

lines(h2$mean, col=3, type="l")

lines(h3$mean, col=5, type="l")

legend("bottomleft", lty=1, col=c(2,3,4,5),c("Holt's Linear", "Additive Damped", "Multiplicative Damped"), cex=0.75)

accuracy(h1, rsv2house)[,c(2,3,5,6)]
accuracy(h2,rsv2house)[,c(2,3,5,6)]
accuracy(h3, rsv2house)[,c(2,3,5,6)]
res <- residuals(h2)

checkresiduals(res)



```


## Exercise 2.4
Generate forecasts using ETS. First select the appropriate model(s) yourself
and discuss their performance. Compare these models with the results of the
automated ETS procedure. Check the residual diagnostics and the forecast
accuracy for the various ETS models you’ve considered

A)Since the data is non seasonal we will use the below model, f1 and f6 are the best model

```{r }

e1 <- ets(rsv1house,"AAN")
e2 <- ets(rsv1house,"MNN")
e3 <- ets(rsv1house,"ANN")
e4 <- ets(rsv1house,"MAN")
e5 <- ets(rsv1house,"MMN")
e6 <- ets(rsv1house, model = "AAN", damped = TRUE)
e7 <- ets(rsv1house, model = "MAN", damped = TRUE)
e8 <- ets(rsv1house, model = "MMN", damped = TRUE)
auto_ets = ets(rsv1house)

auto_ets = ets(rsv1house)
f1 = forecast(e1,h=h)
f2 = forecast(e2,h=h)
f3 = forecast(e3,h=h)
f4 = forecast(e4,h=h)
f5 = forecast(e5,h=h)
f6 = forecast(e6,h=h)
f7 = forecast(e7,h=h)
f8 = forecast(e8,h=h)

accuracy(f1,rsv2house)[,c(2,3,5,6)]
accuracy(f2,rsv2house)[,c(2,3,5,6)]
accuracy(f3,rsv2house)[,c(2,3,5,6)]
accuracy(f4,rsv2house)[,c(2,3,5,6)]
accuracy(f5,rsv2house)[,c(2,3,5,6)]
accuracy(f6,rsv2house)[,c(2,3,5,6)]
accuracy(f7,rsv2house)[,c(2,3,5,6)]
accuracy(f8,rsv2house)[,c(2,3,5,6)]

```





## Exercise 2.5
Generate forecasts using ARIMA. First select the appropriate model(s) yourself
and discuss their performance. Compare these models with the results of the
auto.arima procedure. Check the residual diagnostics and the forecast accuracy
for the various ARIMA models you’ve considered.

```{r }
auto_arima <-auto.arima(rsv1house, seasonal=FALSE, allowdrift = TRUE)
summary(auto_arima)
f1<-forecast(auto_arima,h=h)
accuracy(f1,rsv2house)[,c(2,3,5,6)]
checkresiduals(auto_arima)

best_model <- Arima(rsv1house, order=c(0,2,1))
checkresiduals(best_model)
f2<-forecast(best_model,h=h)
accuracy(f2,rsv2house)[,c(2,3,5,6)]
checkresiduals(best_model)


main_dataset = Arima(rsvhouse, order = c(0,2,1))

for_main = forecast(main_dataset,h=3)
 plot(for_main)
```

## Exercise 2.6
Compare the different models in terms of residual diagnostics, model fit, and
forecast accuracy. Analyse your results and select your final model.

Answer:

best model based on partition was Exponential followed by Arima

Arima

                  RMSE       MAE     MAPE      MASE
Training set  4994.146  2377.229 3.130373 0.5171411
Test set     13871.456 13120.899 6.468137 2.8543134

ETS
                  RMSE       MAE      MAPE      MASE
Training set  5086.783  2444.953  3.346675 0.5318738
Test set     26642.894 24613.080 12.076095 5.3543163 

Exponential
                 RMSE       MAE     MAPE      MASE
Training set  5003.296  2442.777 3.581799 0.5314005
Test set     12687.380 12043.597 5.942417 2.6199576    -best

Naive
                   ME      RMSE       MAE       MPE     MAPE     MASE      ACF1 Theil's U
Training set  4382.756  7267.137  4596.867  6.454523  7.04680 1.000000 0.5758397        NA
Test set     27326.193 29673.556 27326.193 13.400057 13.40006 5.944525 0.5619291  5.365213



## Exercise 2.7
Finally, generate out of sample forecasts up to 2020, based on the complete time
series. Discuss your results

```{r }


plot(forecast(h1, h=10), include = 80)
```












