---
title: "exercis3"
author: "ajay parihar"
date: "April 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



###EXERCISE 3
Q>For this exercise, ﬁnd a recent and relevant time series to forecast, using the techniques that have been discussed during the lectures. In your analysis, set up a carefully selected forecasting process, taking data considerations and implementation issues into account. Describe your approach, and motivate your choices. The data set should be original (not from R packages), recent, suﬀciently long and must include a seasonal component. The data should be analyzed using at least two techniques, and results should be compared.

A> In this exercise we are looking to predict amazon stock price based on historic data. The process covers the following steps:
1)data import
2)creating a time series
3)train/test split
4) jung box testing
5)  arima and tbats functions



```{r amazon stock forecasting }
library(ggplot2)
library(forecast)
library(plotly)
#library(ggfortify)
library(tseries)
library(gridExtra)
#library(docstring)
library(readr)
#library(here)
setwd("C:\\Users\\aparihar\\Downloads")
amazon<-read_csv("HistoricalQuotes.csv")
amazon[is.na(amazon)] <- 0
amazon<-amazon[,-c(3:6)]
amazon$close<-as.numeric(amazon$close)
amazon$date<-as.Date(amazon$date,format="%m/%d/%Y")
ggplot(amazon, aes(date, close)) + geom_line() +
   xlab("test") + ylab("Price")
amazon <- ts(amazon$close, start=c(2009, 5), freq=12)


```


Stationery testing

```{r }
amazon_train <- ts(amazon, start=c(2009, 5), end=c(2015, 5), freq=12)
 Box.test(amazon, lag = 20, type = 'Ljung-Box')
 adf.test(amazon)
 

###Decomposition  
decompose_amazon = decompose(amazon_train, "additive")
 
plot(as.ts(decompose_amazon$seasonal))
plot(as.ts(decompose_amazon$trend))
plot(as.ts(decompose_amazon$random))
plot(decompose_amazon)

acf(amazon_train, 'Amazon')

amz<-auto.arima(amazon_train)
amz
tsDiff <- diff(amazon_train)
plot.ts(tsDiff)
acf(tsDiff, 'S&P 500')

amztff<-auto.arima(tsDiff)

refitamz <- Arima(amazon_train, model=amz)
 accuracy(refitamz)
refittff <- Arima(tsDiff, model=amztff)
 accuracy(refittff)
####since RMSE is lower for amazon_train we will use auto arima amz 
 
fit <- Arima(amazon_train, order = c(0,1,2), include.drift = TRUE) 
summary(fit)

for_amazon_all <- forecast(fit, h = 12)
 plot(for_amazon_all)




m_tbats = tbats(amazon_train)
accuracy(m_tbats)
f_tbats = forecast(m_tbats, h=24)
plot(f_tbats)
```

